---
title: "Plot statistics from VCF"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(ggpubr)
```

- Plotting annotations that GATK suggests to use for hard-filtering. The annotations are QD (QUalByDepth), FS (FisherStrand), SOR (StrandOddsRatio), MQ (RMSMappingQuality), MQRankSum (MappingQualityRankSumTest).

- Plotted is the density plot of each of the annotations. Vertial red lines represent the cutoff value suggested by GATK. 

- This is before any filtering. 

- Regions of interests are chrA (autosomes), chrX, chrY, and chrM.

```{r}
chrA = read.table("c://Users/tuyen/Documents/postdoc_asu/projects/ngs_filtering/out/chrA_prefiltering_annotations.txt", header=T)
chrX = read.table("c://Users/tuyen/Documents/postdoc_asu/projects/ngs_filtering/out/chrX_prefiltering_annotations.txt", header=T)
chrY = read.table("c://Users/tuyen/Documents/postdoc_asu/projects/ngs_filtering/out/chrY_prefiltering_annotations.txt", header=T)
chrM = read.table("c://Users/tuyen/Documents/postdoc_asu/projects/ngs_filtering/out/chrM_prefiltering_annotations.txt", header=T)
```

## QualByDepth (QD)

```{r}
ggplot(data=chrA, aes(x=QD)) +
  geom_density(size=1) +
  geom_vline(xintercept = 2, col="red", size=1) +
  labs(x="QD", y="Density", title="chrA")
ggplot(data=chrX, aes(x=QD)) +
  geom_density(size=1) +
  geom_vline(xintercept = 2, col="red", size=1) +
  labs(x="QD", y="Density", title="chrX")
ggplot(data=chrY, aes(x=QD)) +
  geom_density(size=1) +
  geom_vline(xintercept = 2, col="red", size=1) +
  labs(x="QD", y="Density", title="chrY")
ggplot(data=chrM, aes(x=QD)) +
  geom_density(size=1) +
  geom_vline(xintercept = 2, col="red", size=1) +
  labs(x="QD", y="Density", title="chrM")
```

## FisherStrand (FS)

```{r}
ggplot(data=chrA, aes(x=FS)) +
  geom_density(size=1) +
  geom_vline(xintercept = 60, col="red", size=1) +
  labs(x="FS", y="Density", title="chrA")
ggplot(data=chrX, aes(x=FS)) +
  geom_density(size=1) +
  geom_vline(xintercept = 60, col="red", size=1) +
  labs(x="FS", y="Density", title="chrX")
ggplot(data=chrY, aes(x=FS)) +
  geom_density(size=1) +
  geom_vline(xintercept = 60, col="red", size=1) +
  labs(x="FS", y="Density", title="chrY")
ggplot(data=chrM, aes(x=FS)) +
  geom_density(size=1) +
  geom_vline(xintercept = 60, col="red", size=1) +
  labs(x="FS", y="Density", title="chrM")
```

## StrandOddsRatio (SOR)

```{r}
ggplot(data=chrA, aes(x=SOR)) +
  geom_density(size=1) +
  geom_vline(xintercept = 3, col="red", size=1) +
  labs(x="SOR", y="Density", title="chrA")
ggplot(data=chrX, aes(x=SOR)) +
  geom_density(size=1) +
  geom_vline(xintercept = 3, col="red", size=1) +
  labs(x="SOR", y="Density", title="chrX")
ggplot(data=chrY, aes(x=SOR)) +
  geom_density(size=1) +
  geom_vline(xintercept = 3, col="red", size=1) +
  labs(x="SOR", y="Density", title="chrY")
ggplot(data=chrM, aes(x=SOR)) +
  geom_density(size=1) +
  geom_vline(xintercept = 3, col="red", size=1) +
  labs(x="SOR", y="Density", title="chrM")
```

## RMSMappingQuality (MQ)

```{r}
ggplot(data=chrA, aes(x=MQ)) +
  geom_density(size=1) +
  geom_vline(xintercept = 40, col="red", size=1) +
  labs(x="MQ", y="Density", title="chrA")
ggplot(data=chrX, aes(x=MQ)) +
  geom_density(size=1) +
  geom_vline(xintercept = 40, col="red", size=1) +
  labs(x="MQ", y="Density", title="chrX")
ggplot(data=chrY, aes(x=MQ)) +
  geom_density(size=1) +
  geom_vline(xintercept = 40, col="red", size=1) +
  labs(x="MQ", y="Density", title="chrY")
ggplot(data=chrM, aes(x=MQ)) +
  geom_density(size=1) +
  geom_vline(xintercept = 40, col="red", size=1) +
  labs(x="MQ", y="Density", title="chrM")
```

## MappingQualityRankSumTest (MQRankSum)

```{r}
ggplot(data=chrA, aes(x=MQRankSum)) +
  geom_density(size=1) +
  geom_vline(xintercept = -12.5, col="red", size=1) +
  labs(x="MQRankSum", y="Density", title="chrA")
ggplot(data=chrX, aes(x=MQRankSum)) +
  geom_density(size=1) +
  geom_vline(xintercept = -12.5, col="red", size=1) +
  labs(x="MQRankSum", y="Density", title="chrX")
ggplot(data=chrY, aes(x=MQRankSum)) +
  geom_density(size=1) +
  geom_vline(xintercept = -12.5, col="red", size=1) +
  labs(x="MQRankSum", y="Density", title="chrY")
ggplot(data=chrM, aes(x=MQRankSum)) +
  geom_density(size=1) +
  geom_vline(xintercept = -12.5, col="red", size=1) +
  labs(x="MQRankSum", y="Density", title="chrM")
```

## ReadPosRankSumTest (ReadPosRankSum)

```{r}
ggplot(data=chrA, aes(x=ReadPosRankSum)) +
  geom_density(size=1) +
  geom_vline(xintercept = -8, col="red", size=1) +
  labs(x="ReadPosRankSum", y="Density", title="chrA")
ggplot(data=chrX, aes(x=ReadPosRankSum)) +
  geom_density(size=1) +
  geom_vline(xintercept = -8, col="red", size=1) +
  labs(x="ReadPosRankSum", y="Density", title="chrX")
ggplot(data=chrY, aes(x=ReadPosRankSum)) +
  geom_density(size=1) +
  geom_vline(xintercept = -8, col="red", size=1) +
  labs(x="ReadPosRankSum", y="Density", title="chrY")
ggplot(data=chrM, aes(x=ReadPosRankSum)) +
  geom_density(size=1) +
  geom_vline(xintercept = -8, col="red", size=1) +
  labs(x="ReadPosRankSum", y="Density", title="chrM")
```